Unknown

### 2024-04-10 ms/ms.qmd
revised structure of data analysis section

```{r load-objects, echo=FALSE, eval=TRUE}
library(dplyr)
library(readr)
library(stringr)
library(tibble)

@@

@@

The raw data set consisted of `r prettyNum(aa_stats$n_rh_curve0, big.mark = ",")` \agcurve{} curves with an average of `r formatC(aa_stats$n_point_per_rh_curve0, digits = 3)` points per curve. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into six sequential steps \autoref{tbl-cleaning1}.

@@

The raw data set consisted of `r prettyNum(aa_stats$n_aq_curve0, big.mark = ",")` \aqcurve{} curves with an average of `r formatC(aa_stats$n_point_per_aq_curve0, digits = 3)`. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into two sequential steps \autoref{tbl-cleaning2}.

@@

### Data analysis

#### General points (not sure where this goes yet)

- models fit in Stan/brms
- we chose the number of chains, warmup and sampling iterations, and maximum treedepth so that parameter estimates converged ($\hat{R} < 1.01$ [@gelman_inference_1992]) and the effective sample size (ESS) for each parameter was $> 10^3$.

We fit Bayesian models with MCMC sampling in the probabilistic programming language *Stan* [@stan_development_team_stan_2023]. We used CmdStan version `r fit_aa$version$cmdstan` and **cmdstanr** version `r fit_aa$version$cmdstanr` [@gabry_cmdstanr_2023] to interface with *R* version `r getRversion()` [@r_core_team_r:_2023]. We sampled the posterior distribution from `r dim_fit_aa[2]` chains with `r dim_fit_aa[1]` iterations each after `r dim_fit_aa[1]` warmup iterations per chain. We estimated parameters and confidence intervals as the median and 95% quantile intervals of the posterior, respectively. 
#### *Stan* models

- high level overview of differences among models \autoref{tbl-models}
- next section describes common elements of all models and additional details about parameters

@@

#### Parameter estimation

>>>>

```{r load-objects, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(stringr)
library(tibble)

source("../r/functions.R")

@@

### Cleaning AQ curves

### Bayesian data analysis
n_model = 5 |>
  number_to_word()

@@

The raw data set consisted of `r prettyNum(aa_stats$n_rh_curve0, big.mark = ",")` \agcurve{} curves with an average of `r formatC(aa_stats$n_point_per_rh_curve0, digits = 3)` points per curve. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into six sequential steps (\autoref{tbl-cleaning1}).

@@

The raw data set consisted of `r prettyNum(aa_stats$n_aq_curve0, big.mark = ",")` \aqcurve{} curves with an average of `r formatC(aa_stats$n_point_per_aq_curve0, digits = 3)`. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into two sequential steps (\autoref{tbl-cleaning2}).

@@

### Bayesian data analysis in *Stan*

We fit `r n_model` models to test predictions of competing hypotheses about why amphistomy advantage (\aax) might be greater for leaves in sunny, open habitats. This section provides an overview of differences among models (\autoref{tbl-models}). The next sections describe how we fit models in *Stan*, all model parameters and priors, and specific predictions about parameter values for each hypothesis.

@@

#### Fitting models in *Stan*

We fit Bayesian models with MCMC sampling in the probabilistic programming language *Stan* [@stan_development_team_stan_2023]. We used CmdStan version `r fit_aa$version$cmdstan` and **cmdstanr** version `r fit_aa$version$cmdstanr` [@gabry_cmdstanr_2023] to interface with *R* version `r getRversion()` [@r_core_team_r:_2023]. We sampled the posterior distribution from `r dim_fit_aa[2]` chains with `r dim_fit_aa[1]` iterations each after `r dim_fit_aa[1]` warmup iterations per chain. We estimated parameters and confidence intervals as the median and 95% quantile intervals of the posterior, respectively. We chose the number of chains, warmup and sampling iterations, and maximum treedepth so that parameter estimates converged ($\hat{R} < 1.01$ [@gelman_inference_1992]) and the effective sample size (ESS) for each parameter was $> 10^3$.

#### Parameter estimation and priors

### 2024-04-10 ms/ms.tex
revised structure of data analysis section

\begin{verbatim}

Attaching package: 'dplyr'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:stats':

    filter, lag
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    intersect, setdiff, setequal, union
\end{verbatim}

@@

The raw data set consisted of 2,370 \agcurve{} curves with an average of
63.2 points per curve. Manual curation of a data set this size in a
principled, consistent manner is not feasible. Therefore, we automated
data cleaning using custom \emph{R} scripts. Cleaning is divided into
six sequential steps \autoref{tbl-cleaning1}.

@@

The raw data set consisted of 658 \aqcurve{} curves with an average of
19.4. Manual curation of a data set this size in a principled,
consistent manner is not feasible. Therefore, we automated data cleaning
using custom \emph{R} scripts. Cleaning is divided into two sequential
steps \autoref{tbl-cleaning2}.

@@

\subsubsection{Data analysis}\label{data-analysis}

\paragraph{General points (not sure where this goes
yet)}\label{general-points-not-sure-where-this-goes-yet}

\begin{itemize}
\tightlist
\item
  models fit in Stan/brms
\item
  we chose the number of chains, warmup and sampling iterations, and
  maximum treedepth so that parameter estimates converged
  (\(\hat{R} < 1.01\) (\emph{10})) and the effective sample size (ESS)
  for each parameter was \(> 10^3\).
\end{itemize}

We fit Bayesian models with MCMC sampling in the probabilistic
programming language \emph{Stan} (\emph{11}). We used CmdStan version
2.33.1 and \textbf{cmdstanr} version 0.7.1 (\emph{12}) to interface with
\emph{R} version 4.3.2 (\emph{13}). We sampled the posterior
distribution from 4 chains with 1000 iterations each after 1000 warmup
iterations per chain. We estimated parameters and confidence intervals
as the median and 95\% quantile intervals of the posterior,
respectively. \#\#\#\# \emph{Stan} models

\begin{itemize}
\tightlist
\item
  high level overview of differences among models \autoref{tbl-models}
\item
  next section describes common elements of all models and additional
  details about parameters
\end{itemize}

@@

\paragraph{Parameter estimation}\label{parameter-estimation}

@@

\hspace{1em}\cellcolor{gray!10}{$\beta_{\mathrm{AA},\text{PPFD},0}$} & \cellcolor{gray!10}{intercept of accession-level \aax{} when native \ppfdequals{0}}\\
\hspace{1em}$\beta_{\mathrm{AA},\text{PPFD},1}$ & slope of native \ppfd{} on accession-level \aax{}\\

@@

We tested for effects of light intensity, light treatment, accession,
and their interactions on \aax. All models included effects of light
intensity and light treatment, as well as random effects of accession
and replicate within accession. More complex models included
interactions between light intensity and light treatment, as well as
random effects of accession on the effects of light intensity and light
treatment. We used a weakly informative normal prior with mean \(0\) and
standard deviation \(10\) for fixed effects light intensity and
treatment. We used a weakly informative normal prior with mean \(-3\)
and standard deviation \(5\) for the the random effect standard
deviation of replicate within accession. We accounted for the
phylogenetic structure among the random effects of accession using an
Ornstein-Uhlenbeck (OU) process (\textbf{hansen\_stabilizing\_1997?}).
The expected covariance between accessions \(i\) and \(j\)
(\(\text{Cov}(i, j)\) ) is:

@@

We tested competing predictions through a combination of model selection
and parameter estimation from the posterior of selected models. We
started with a base model that includes fixed effects of light intensity
and light treatment on \aax, a random effect of accession on \aax, and a
linear regression of PAR on accession-level \aax. We next compared the
fit of the base model to a more complex with an interaction between
light intensity and light treatment using LOOIC with the \textit{R}
package \textbf{loo} version 2.7.0 (\emph{14}). These comparisons
allowed us to test the main predictions of each hypothesis and for a
reinforcing interactions between assimilatory and plastic hypotheses
(tab:predictions). However, these models cannot test predictions when
there is an interaction with the constitutive hypothesis, because this
hypothesis posits accession-levels variation in responses to light
intensity and/or light treatment. Therefore, we tested for significant
variation among accessions in light intensity and light treatment by
comparing LOOIC values of models with and without random interactions
between these factors and accession.

@@

\bibitem[\citeproctext]{ref-gelman_inference_1992}
\CSLLeftMargin{10. }%
\CSLRightInline{A. Gelman, D. B. Rubin, Inference from iterative
simulation using multiple sequences. \emph{Statistical Science}
\textbf{7}, 457--472 (1992).}

\bibitem[\citeproctext]{ref-stan_development_team_stan_2023}
\CSLLeftMargin{11. }%
\CSLRightInline{Stan Development Team, \emph{Stan {Modeling} {Language}
{Users} {Guide} and {Reference} {Manual}} (2023;
\url{https://mc-stan.org}).}

\bibitem[\citeproctext]{ref-gabry_cmdstanr_2023}
\CSLLeftMargin{12. }%
\CSLRightInline{J. Gabry, R. Češnovar, \emph{Cmdstanr: {R} {Interface}
to '{CmdStan}'} (2023;
\href{https://mc-stan.org/cmdstanr,\%20https://discourse.mc-stan.org}{https://mc-stan.org/cmdstanr,
https://discourse.mc-stan.org}).}

\bibitem[\citeproctext]{ref-r_core_team_r:_2023}
\CSLLeftMargin{13. }%
\CSLRightInline{R Core Team, \emph{R: {A} {Language} and {Environment}
for {Statistical} {Computing}} (R Foundation for Statistical Computing,
Vienna, Austria, 2023; \url{http://www.R-project.org/}).}

\bibitem[\citeproctext]{ref-vehtari_practical_2017}
\CSLLeftMargin{14. }%
\CSLRightInline{A. Vehtari, A. Gelman, J. Gabry,
\href{https://doi.org/10.1007/s11222-016-9696-4}{Practical {Bayesian}
model evaluation using leave-one-out cross-validation and {WAIC}}.
\emph{Statistics and Computing} \textbf{27}, 1413--1432 (2017).}

>>>>

@@

The raw data set consisted of 2,370 \agcurve{} curves with an average of
63.2 points per curve. Manual curation of a data set this size in a
principled, consistent manner is not feasible. Therefore, we automated
data cleaning using custom \emph{R} scripts. Cleaning is divided into
six sequential steps (\autoref{tbl-cleaning1}).

@@

The raw data set consisted of 658 \aqcurve{} curves with an average of
19.4. Manual curation of a data set this size in a principled,
consistent manner is not feasible. Therefore, we automated data cleaning
using custom \emph{R} scripts. Cleaning is divided into two sequential
steps (\autoref{tbl-cleaning2}).

@@

\subsubsection{\texorpdfstring{Bayesian data analysis in
\emph{Stan}}{Bayesian data analysis in Stan}}\label{bayesian-data-analysis-in-stan}

We fit five models to test predictions of competing hypotheses about why
amphistomy advantage (\aax) might be greater for leaves in sunny, open
habitats. This section provides an overview of differences among models
(\autoref{tbl-models}). The next sections describe how we fit models in
\emph{Stan}, all model parameters and priors, and specific predictions
about parameter values for each hypothesis.

@@

\paragraph{\texorpdfstring{Fitting models in
\emph{Stan}}{Fitting models in Stan}}\label{fitting-models-in-stan}

We fit Bayesian models with MCMC sampling in the probabilistic
programming language \emph{Stan} (\emph{10}). We used CmdStan version
2.33.1 and \textbf{cmdstanr} version 0.7.1 (\emph{11}) to interface with
\emph{R} version 4.3.2 (\emph{12}). We sampled the posterior
distribution from 4 chains with 1000 iterations each after 1000 warmup
iterations per chain. We estimated parameters and confidence intervals
as the median and 95\% quantile intervals of the posterior,
respectively. We chose the number of chains, warmup and sampling
iterations, and maximum treedepth so that parameter estimates converged
(\(\hat{R} < 1.01\) (\emph{13})) and the effective sample size (ESS) for
each parameter was \(> 10^3\).

\paragraph{Parameter estimation and
priors}\label{parameter-estimation-and-priors}

@@

\hspace{1em}\hspace{1em}\cellcolor{gray!10}{$\beta_{\mathrm{AA},\text{PPFD},0}$} & \cellcolor{gray!10}{intercept of accession-level \aax{} when native \ppfdequals{0}}\\
\hspace{1em}\hspace{1em}$\beta_{\mathrm{AA},\text{PPFD},1}$ & slope of native \ppfd{} on accession-level \aax{}\\
\hspace{1em}\cellcolor{gray!10}{$\alpha_{\mathrm{AA},\text{PPFD}}$} & \cellcolor{gray!10}{decay rate of phylogenetic covariance in residuals of model testing effect of native \ppfd{} on accession-level \aax}\\
\hspace{1em}$\sigma^2_{\mathrm{AA},\text{PPFD}}$ & phylogenetic diffusion rate in residuals of model testing effect of native \ppfd{} on accession-level \aax\\

@@

We tested for effects of light intensity, light treatment, accession,
and their interactions on \aax. All models included effects of light
intensity and light treatment, as well as random effects of accession
and replicate within accession. More complex models included
interactions between light intensity and light treatment, as well as
random effects of accession on the effects of light intensity and light
treatment. We used a weakly informative normal prior with mean \(0\) and
standard deviation \(10\) for fixed effects light intensity and
treatment. We used a weakly informative normal prior with mean \(-3\)
and standard deviation \(5\) for the the random effect standard
deviation of replicate within accession. We accounted for the
phylogenetic structure among the random effects of accession using an
Ornstein-Uhlenbeck (OU) process (\emph{14}). The expected covariance
between accessions \(i\) and \(j\) (\(\text{Cov}(i, j)\) ) is:

@@

We tested competing predictions through a combination of model selection
and parameter estimation from the posterior of selected models. We
started with a base model that includes fixed effects of light intensity
and light treatment on \aax, a random effect of accession on \aax, and a
linear regression of PAR on accession-level \aax. We next compared the
fit of the base model to a more complex with an interaction between
light intensity and light treatment using LOOIC with the \textit{R}
package \textbf{loo} version 2.7.0 (\emph{15}). These comparisons
allowed us to test the main predictions of each hypothesis and for a
reinforcing interactions between assimilatory and plastic hypotheses
(tab:predictions). However, these models cannot test predictions when
there is an interaction with the constitutive hypothesis, because this
hypothesis posits accession-levels variation in responses to light
intensity and/or light treatment. Therefore, we tested for significant
variation among accessions in light intensity and light treatment by
comparing LOOIC values of models with and without random interactions
between these factors and accession.

@@

\bibitem[\citeproctext]{ref-stan_development_team_stan_2023}
\CSLLeftMargin{10. }%
\CSLRightInline{Stan Development Team, \emph{Stan {Modeling} {Language}
{Users} {Guide} and {Reference} {Manual}} (2023;
\url{https://mc-stan.org}).}

\bibitem[\citeproctext]{ref-gabry_cmdstanr_2023}
\CSLLeftMargin{11. }%
\CSLRightInline{J. Gabry, R. Češnovar, \emph{Cmdstanr: {R} {Interface}
to '{CmdStan}'} (2023;
\href{https://mc-stan.org/cmdstanr,\%20https://discourse.mc-stan.org}{https://mc-stan.org/cmdstanr,
https://discourse.mc-stan.org}).}

\bibitem[\citeproctext]{ref-r_core_team_r:_2023}
\CSLLeftMargin{12. }%
\CSLRightInline{R Core Team, \emph{R: {A} {Language} and {Environment}
for {Statistical} {Computing}} (R Foundation for Statistical Computing,
Vienna, Austria, 2023; \url{http://www.R-project.org/}).}

\bibitem[\citeproctext]{ref-gelman_inference_1992}
\CSLLeftMargin{13. }%
\CSLRightInline{A. Gelman, D. B. Rubin, Inference from iterative
simulation using multiple sequences. \emph{Statistical Science}
\textbf{7}, 457--472 (1992).}

\bibitem[\citeproctext]{ref-hansen_stabilizing_1997}
\CSLLeftMargin{14. }%
\CSLRightInline{T. F. Hansen,
\href{https://doi.org/10.1111/j.1558-5646.1997.tb01457.x}{Stabilizing
selection and the comparative analysis of adaptation}. \emph{Evolution}
\textbf{51}, 1341--1351 (1997).}

\bibitem[\citeproctext]{ref-vehtari_practical_2017}
\CSLLeftMargin{15. }%
\CSLRightInline{A. Vehtari, A. Gelman, J. Gabry,
\href{https://doi.org/10.1007/s11222-016-9696-4}{Practical {Bayesian}
model evaluation using leave-one-out cross-validation and {WAIC}}.
\emph{Statistics and Computing} \textbf{27}, 1413--1432 (2017).}
